---
title: "Pair and dropout communities SynCom"
subtitle: "Thoenen et al. BX SynComs"
author: "Lisa Thoenen"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    code_folding: show
editor_options: 
  chunk_output_type: console
---

This script provides the analysis for the pair and dropout metabolite data which were used to create Figure 2.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
rm(list = ls())
```

```{r,include=FALSE}
# install.packages("dplyr") 
# install.packages("magrittr") # for the %>% function
# install.packages("ggplot2") # plots
# install.packages("readxl") # to read excels
# install.packages("stringr")
# install.packages("MESS") # for auc function
# install.packages("purrr")
# install.packages("multcomp")
# install.packages("emmeans")
# install.packages("ggpubr")
# install.packages("forcats")
# install.packages("readr")
# install.packages("ggpmisc")
# install.packages("pander")
# install.packages("rstatix")
# install.packages("pheatmap")
```

```{r,include=FALSE}
library("dplyr")
library("magrittr")
library("ggplot2")
library("readxl")
library("tidyr")
library("readr")
library("stringr")
library("MESS") 
library("purrr")
library("multcomp")
library("emmeans")
library("ggpubr")
library("forcats")
library("readr")
library("ggpmisc")
library("pander")
library("rstatix")
library("pheatmap")
```

# Data
```{r, echo=FALSE, warning=FALSE}
# Metadata
metadata <- read_excel("./Input/3_samples_metabolites_combo.xlsx") %>% dplyr::select(-random, -Nr) %>% distinct() %>%  
  mutate(Sample_ID = gsub("SCR", "metSC", Sample_ID)) %>% 
  mutate(Strain = gsub("SCR", "metSC", Strain)) %>% 
  mutate(Sample_ID = gsub("SCS", "nonSC", Sample_ID)) %>% 
  mutate(Strain = gsub("SCS", "nonSC", Strain)) 
# Metabolite data
met <- read_excel("./Input/3_Bact_combo_metabolites.xlsx") %>% 
  mutate(Name = gsub("SCR", "metSC", Name)) %>% 
  mutate(Name = gsub("SCS", "nonSC", Name)) %>% 
  dplyr::rename(Std_conc = `Std. Conc`)
# Molecular weights
mol <- read_excel("./Input/Molecular_weight_BXDs.xlsx")
```

## Benzoxazinoids quantification

```{r, echo=FALSE, warning=FALSE}
met_Std <- met %>% 
  filter(Type %in% "Standard") %>% 
  filter(Std_conc != "0") %>% 
  filter(Compound != c("AAPO", "MHPA")) %>% 
  filter(!Name %in% c("20211123_MH_LT_Bact_MHPA_40ng", "20211123_MH_LT_Bact_MHPA_200ng", "20211123_MH_LT_Bact_MHPA_1000ng")) %>% 
  dplyr::select(Std_conc, Area, Compound) %>% 
  filter(!Std_conc %in% "NA")
```

```{r, echo=FALSE, warning=FALSE}
met_Std %>% as.data.frame() %>% 
  filter(!Compound %in% c("AAPO", "MHPA")) %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>% 
  mutate(Area = as.numeric(Area)) %>% 
  ggplot(aes(x= Std_conc, y = Area)) + 
  geom_point(aes(colour = Compound), show.legend = FALSE) +
  geom_smooth(method = "lm", color="black", formula = y ~ 0 + x) +
  stat_poly_eq(formula = y ~ 0 + x,
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                parse = TRUE, label.y.npc = "top", label.x.npc = "left") +
  stat_fit_glance(method = "lm", method.args = list(formula = y ~ 0 + x),
                      geom = 'text', aes(label = paste("P-val. = ", 
                      signif(..p.value.., digits = 4), sep = "")), label.y = "bottom", label.x = "center") +
  facet_wrap(~Compound, scales = "free") +
  labs(y = "Area",
       x = "Concentration")
```

```{r, warning=F, echo=F, message=F}
St_slopes <- met_Std %>% 
  filter(!Compound %in% c("AAPO", "MHPA")) %>% 
  mutate(Compound = as.factor(Compound)) %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>%
  group_by(Compound) %>% 
  do(lm(Area ~ 0 + Std_conc,  data = .) %>%  coef() %>%   as.data.frame() %>% slice(1))

colnames(St_slopes) <- c("Compound", "Slope")
```

```{r, echo=FALSE, warning=FALSE}
met_Std_MHPA <- met %>% 
  filter(Type %in% "Standard") %>% 
  filter(Std_conc != "0") %>% 
  filter(Compound %in% "MHPA") %>% 
  filter(Name %in% c("20211123_MH_LT_Bact_MHPA_40ng", "20211123_MH_LT_Bact_MHPA_200ng", "20211123_MH_LT_Bact_MHPA_1000ng")) %>% 
  dplyr::select(Std_conc, Area, Compound) %>% 
  filter(!Std_conc %in% "NA")
```

```{r, echo=FALSE, warning=FALSE}
met_Std_MHPA %>% as.data.frame() %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>% 
  mutate(Area = as.numeric(Area)) %>% 
  ggplot(aes(x= Std_conc, y = Area)) + 
  geom_point(aes(colour = Compound), show.legend = FALSE) +
  geom_smooth(method = "lm", color="black", formula = y ~ 0 + x) +
  stat_poly_eq(formula = y ~ 0 + x,
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                parse = TRUE, label.y.npc = "top", label.x.npc = "left") +
  stat_fit_glance(method = "lm", method.args = list(formula = y ~ 0 + x),
                      geom = 'text', aes(label = paste("P-val. = ", 
                      signif(..p.value.., digits = 4), sep = "")), label.y = "bottom", label.x = "center") +
  facet_wrap(~Compound, scales = "free") +
  labs(y = "Area",
       x = "Concentration")
```

```{r, warning=F, echo=F, message=F}
St_slopes_MHPA <- met_Std_MHPA %>% 
  mutate(Compound = as.factor(Compound)) %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>%
  group_by(Compound) %>% 
  do(lm(Area ~ 0 + Std_conc,  data = .) %>%  coef() %>%   as.data.frame() %>% slice(1))
```

```{r, echo=FALSE, warning=FALSE}
colnames(St_slopes_MHPA) <- c("Compound", "Slope")
```

```{r, echo=FALSE, warning=FALSE}
St_slopes <- rbind(St_slopes, St_slopes_MHPA)
```

## Data formatting
```{r, echo=FALSE, warning=FALSE}
met$Name <- gsub("20211123_MH_LT_Bact_", "", met$Name)
met$Name <- gsub("LRC7-O", "LRC7.O", met$Name)
met %<>% filter(Type %in% "Analyte")
met <- left_join(met, mol, by = "Compound")
met %<>% mutate(M = as.numeric(M)) %>% mutate(Name = gsub("NBC_D", "NBC_c_D", Name)) %>% mutate(Name = gsub("NBC_M", "NBC_c_M", Name))
```

```{r, echo=FALSE, warning=FALSE}
metadata %<>% dplyr::rename(Name = Sample_ID) 
met <- left_join(met, metadata, by = "Name") 
met %<>% as.data.frame()
```

```{r, echo=FALSE, warning=FALSE}
met$Treatment <- factor(met$Treatment, levels = c("DMSO", "MBOA"))
met$Strain <- gsub("_d", "", met$Strain)
met$Strain <- gsub("_c", "", met$Strain)

met$Strain_exp <- interaction(met$Strain, met$experiment, sep = "_")
met$Strain_exp <- gsub("_dropout", "_d", met$Strain_exp)
met$Strain_exp <- gsub("_combo", "_c", met$Strain_exp)

met <- left_join(met, St_slopes, by = "Compound")
```

```{r, echo=FALSE, warning=FALSE}
met$Area_sample <- as.numeric(met$Area) / 0.02 
met$ng_sample <- met$Area_sample / met$Slope 
met$uMol <- (met$ng_sample) / met$M
```

```{r, echo=FALSE, warning=FALSE}
met %<>% mutate(Compound = gsub("MHPA", "HMPAA", Compound)) %>% 
         mutate(Compound = gsub("MAPH", "AMP", Compound))
```

```{r, echo=FALSE, warning=FALSE}
met$cols <- as.character(met$Compound)
met[met$Compound == "AMPO", ]$cols <- "darkred"
met[met$Compound == "AAMPO", ]$cols <- "firebrick1"
met[met$Compound == "HMPAA", ]$cols <- "gold2"
met[met$Compound == "MBOA", ]$cols <- "darkblue"

met[met$Compound == "HMBOA", ]$cols <- "skyblue"
met[met$Compound == "BOA", ]$cols <- "slateblue"

met[met$Compound == "MBOA-Glc", ]$cols <- "slategray"
met[met$Compound == "HMPMA", ]$cols <-  "slategray"
met[met$Compound == "DIMBOA", ]$cols <-  "slategray"
met[met$Compound == "APO", ]$cols <-  "slategray"
met[met$Compound == "AAPO", ]$cols <-  "slategray"


temp <- data.frame(met$Compound, met$cols)
temp <- plyr::ddply(temp, .variables="met.Compound", .fun=unique)   #library(plyr)
level_cols_Compound <- as.character(temp[,2])
names(level_cols_Compound) <- temp[,1]
```

```{r, echo=FALSE, warning=FALSE}
met %<>% filter(!Compound %in% c("VAA", "DIBOA-Glc", "HBOA-Glc")) %>% 
         filter(Strain %in% c("LMB2", "LMI1x", "LMX9", "LMX9231", "LBA112", "LMG1", "LMN1", "LST17", "LBA21", "metSC", "nonSC", "SC3", "SC32", "NBC"))
```

Pairs
- LMB2 + LBA21
- LMB2 pure
- LMB2 + LMG1
- LMB2 + LMN1
- LMB2 + LMX9
- LMB2 + LMX92
- LMB2 + LMX922
- LMB2 + LMX923
- LMB2 + LMX9231
- LMB2 + LST17

Dropout
- complete community - metSC: LMB2, LBA21, LMG1, LMN1, LST17, LMX9, LMX9231
- for each dropout community one strain is left out

SynComs
- metSC: LMB2, LBA21, LMG1, LMN1, LST17, LMX9, LMX9231
- metSC2: LMB2, LBA21, LMG1, LMN1, LST17, LMX9, LMX922

reduced SynComs
- SC3: LMB2, LMX9, LMX9231
- SC32: LMB2, LMX9, LMX922

```{r, echo=FALSE, warning=FALSE}
met %>% filter(Strain %in% "NBC") %>% filter(Compound %in% "MBOA")
```

# Plottiing
## quantitative data

```{r, message=FALSE, echo=FALSE, error=FALSE, warning = FALSE, fig.dim = c(8,8)}
met %>% 
  mutate(Compound = as.factor(Compound)) %>% 
  ggplot(aes(x = Strain, y = as.numeric(Area), group = Treatment)) +
  geom_point(aes(colour = Strain), show.legend = TRUE) +
  facet_wrap(~ Compound, scales = "free") +
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) 
```

```{r, message=FALSE, echo=FALSE, error=FALSE, warning = FALSE, fig.dim = c(8,8)}
met %>% 
  as.data.frame() %>% 
  filter(Treatment %in% "MBOA") %>% 
  filter(!Strain %in% c("LMB2", "NBC")) %>% 
  filter(experiment %in% "combo") %>% 
  mutate(Area = as.numeric(Area)) %>% 
  filter(experiment !="NA") %>% 
  ggplot(aes(x = as.factor(Strain), y = uMol)) +
  geom_bar(aes(fill = Compound), stat = "identity", size = 2, show.legend = TRUE) +
  ylim(0, 600) +
  scale_fill_manual(values = level_cols_Compound)+
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) +
  labs(title = "pairs", 
       x = "")
```


MBOA degradation in paired communities of LMB2 and other strains
```{r, message=FALSE, echo=FALSE, error=FALSE, warning = FALSE, fig.dim = c(8,8)}
pairs_MBOA_met_AAMPO <- met %>% 
  as.data.frame() %>% 
  filter(Compound %in% c("MBOA", "HMPAA", "AMPO", "AAMPO")) %>% 
  mutate(Compound = factor(Compound, levels = c("HMPAA","AAMPO" , "AMPO",  "MBOA"))) %>% 
  filter(Treatment %in% "MBOA") %>% 
  filter(Strain_exp %in% c("NBC_d", "metSC_d", "LST17_c", "LMX9231_c", "LMX9_c", "LBA21_c", "LMN1_c", "LMG1_c", "LMB2_c")) %>% 
  mutate(Strain = factor(Strain, levels = rev(c("NBC", "LMB2", "LBA21", "LMG1", "LMN1", "LMX9", "LMX9231", "LST17", "metSC")))) %>% 
  mutate(Area = as.numeric(Area)) %>% 
  filter(experiment !="NA") %>% 
  ggplot(aes(x = as.factor(Strain), y = uMol)) +
  geom_bar(aes(fill = Compound), stat = "identity", size = 2, show.legend = TRUE) +
  scale_fill_manual(values = c("gold2", "firebrick1","darkred", "darkblue")) +
  theme_classic() +
  theme(axis.text.x=element_text(angle = 0, hjust = 0, vjust = 0.5)) +
  coord_flip()+
  scale_x_discrete(labels= rev(c("NBC", "LMB2", "LBA21 + LMB2", "LMG1 + LMB2", "LMN1 + LMB2", "LMX9 + LMB2", "LMX9231 + LMB2", "LST17 + LMB2", "metSC")))+
  scale_y_continuous(expand = expansion(c(0, 0.1)))+
  labs(title = "pairs", 
       y = "Concentration [µM]",
       x = "",
       fill = "Compound")

pairs_MBOA_met_AAMPO
pairs_MBOA_met_AAMPO_export <- pairs_MBOA_met_AAMPO +
    scale_x_discrete(labels= rev(c("NBC", "LMB2", "LBA21", "LMG1", "LMN1", "LMX9", "LMX9231", "LST17", "metSC")))
  

ggsave(plot = pairs_MBOA_met_AAMPO_export, filename = "Fig_2_pairs_MBOA_met_withAAMPO.svg", path = "Output", width = 8.8, height = 9, dpi = 300, scale = 1, units = "cm", create.dir = TRUE)
```


```{r, message=FALSE,echo=FALSE, error=FALSE, warning = FALSE, fig.dim = c(8,8),  include=F}
#
pairs_MBOA_met <- met %>% 
  as.data.frame() %>% 
  filter(Compound %in% c("MBOA", "HMPAA", "AMPO", "AAMPO")) %>% 
  mutate(Compound = factor(Compound, levels = c("HMPAA","AAMPO" , "AMPO",  "MBOA"))) %>% 
  filter(Treatment %in% "MBOA") %>% 
  filter(Strain_exp %in% c("NBC_d", "metSC_d", "LST17_c", "LMX9231_c", "LMX9_c", "LBA21_c", "LMN1_c", "LMG1_c", "LMB2_c")) %>% 
  mutate(Strain = factor(Strain, levels = rev(c("NBC", "LMB2", "LBA21", "LMG1", "LMN1", "LMX9", "LMX9231", "LST17", "metSC")))) %>% 
  mutate(Area = as.numeric(Area)) %>% 
  filter(experiment !="NA") %>% 
  ggplot(aes(x = as.factor(Strain), y = uMol)) +
  geom_bar(aes(fill = Compound), stat = "identity", size = 2, show.legend = TRUE) +
  scale_fill_manual(values = c("gold2", "firebrick1","darkred", "darkblue")) +
  theme_classic() +
  theme(axis.text.x=element_text(angle = 0, hjust = 0, vjust = 0.5)) +
  coord_flip()+
  scale_x_discrete(labels= rev(c("NBC", "LMB2", "LBA21 + LMB2", "LMG1 + LMB2", "LMN1 + LMB2", "LMX9 + LMB2", "LMX9231 + LMB2", "LST17 + LMB2", "metSC")))+
  scale_y_continuous(expand = expansion(c(0, 0.1)))+
  labs(title = "pairs", 
       y = "Concentration [µM]",
       x = "",
       fill = "Compound")

pairs_MBOA_met
pairs_MBOA_met_export <- pairs_MBOA_met+
    scale_x_discrete(labels= rev(c("NBC", "LMB2", "LBA21", "LMG1", "LMN1", "LMX9", "LMX9231", "LST17", "metSC")))
  

# ggsave(plot = pairs_MBOA_met_export, filename = "Fig_2_pairs_MBOA_met.svg", path = "Output", width = 8.8, height = 9, dpi = 300, scale = 1, units = "cm")
```

Deradation of MBOA in metSC dropout communities
```{r, message=FALSE, echo=FALSE, error=FALSE, warning = FALSE, fig.dim = c(8,8)}
dropout_MBOA_met_AAMPO <- met %>% 
  as.data.frame() %>% 
  filter(Compound %in% c("MBOA", "HMPAA", "AMPO","AAMPO")) %>% 
  mutate(Compound = factor(Compound, levels = c("HMPAA","AAMPO",  "AMPO", "MBOA"))) %>% 
  filter(Treatment %in% "MBOA") %>% 
  filter(Strain_exp %in% c("LBA21_d", "LMN1_d", "LMX9231_d", "LMX9_d", "LMG1_d", "LST17_d", "LMB2_d", "metSC_d", "NBC_d")) %>% 
  mutate(Strain = factor(Strain, levels = rev(c("NBC", "LMB2", "LBA21", "LMG1", "LMN1", "LMX9", "LMX9231", "LST17", "metSC")))) %>% 
  filter(experiment !="NA") %>% 
  unique() %>% 
  ggplot(aes(x = as.factor(Strain), y = uMol)) +
  geom_bar(aes(fill = Compound), stat = "identity", size = 2, show.legend = TRUE) +
  scale_fill_manual(values = c("gold2", "firebrick1","darkred", "darkblue")) +
  coord_flip()+
  scale_x_discrete(labels= rev(c("NBC", "metSC - LMB2", "metSC - LBA21", "metSC - LMG1", "metSC - LMN1", "metSC - LMX9", "metSC - LMX9231", "metSC - LST17", "metSC")))+
  theme_classic() +
  theme(axis.text.x=element_text(angle = 0, hjust = 0, vjust = 0.5)) +
  scale_y_continuous(expand = expansion(c(0, 0.1)))+
  # facet_wrap(~experiment) + 
  labs(title = "dropout", 
       x = "",
       fill = "Compound",
       y = "Concentration [µM]")

dropout_MBOA_met_AAMPO
dropout_MBOA_met_AAMPO_export <- dropout_MBOA_met_AAMPO +
  scale_x_discrete(labels= rev(c("NBC", "LMB2", "LBA21", "LMG1", "LMN1", "LMX9", "LMX9231", "LST17", "metSC")))

ggsave(plot = dropout_MBOA_met_AAMPO_export, filename = "Fig_2_dropout_MBOA_met_withAAMPO.svg", path = "Output", width = 8.8, height = 9, dpi = 300, scale = 1, units = "cm", create.dir = TRUE)
```

Export of data
```{r, message=FALSE, echo=FALSE, error=FALSE, warning = FALSE, fig.dim = c(8,8)}
combined_figure <- ggarrange(dropout_MBOA_met_AAMPO_export, pairs_MBOA_met_AAMPO_export, ncol = 2, common.legend = T, align = "h", legend = "right", labels = c("A", "B"))
combined_figure

ggsave(plot = combined_figure, filename = "Fig_2_combined_met_withAAMPO.svg", path = "Output", width = 16, height = 9, dpi = 300, scale = 1, units = "cm")
ggsave(plot = combined_figure, filename = "Fig_2_combined_met_withAAMPO.pdf", path = "Output", width = 16, height = 9, dpi = 300, scale = 1, units = "cm")
```


```{r, message=FALSE,  include = F}
#echo=FALSE, error=FALSE, warning = FALSE, fig.dim = c(8,8),
#Here without showing AAMPO to focus on MBOA, AMPO and HMPAA
dropout_MBOA_met<- met %>%
  as.data.frame() %>%
  filter(Compound %in% c("MBOA", "HMPAA", "AMPO")) %>%
  mutate(Compound = factor(Compound, levels = c("HMPAA","AMPO", "MBOA"))) %>%
  filter(Treatment %in% "MBOA") %>%
  filter(Strain_exp %in% c("LBA21_d", "LMN1_d", "LMX9231_d", "LMX9_d", "LMG1_d", "LST17_d", "LMB2_d", "metSC_d", "NBC_d")) %>%
  mutate(Strain = factor(Strain, levels = rev(c("NBC", "LMB2", "LBA21", "LMG1", "LMN1", "LMX9", "LMX9231", "LST17", "metSC")))) %>%
  filter(experiment !="NA") %>%
  unique() %>%
  ggplot(aes(x = as.factor(Strain), y = uMol)) +
  geom_bar(aes(fill = Compound), stat = "identity", size = 2, show.legend = TRUE) +
  scale_fill_manual(values = c("gold2", "darkred", "darkblue")) +
  coord_flip()+
  scale_x_discrete(labels= rev(c("NBC", "metSC - LMB2", "metSC - LBA21", "metSC - LMG1", "metSC - LMN1", "metSC - LMX9", "metSC - LMX9231", "metSC - LST17", "metSC")))+
  theme_classic() +
  theme(axis.text.x=element_text(angle = 0, hjust = 0, vjust = 0.5)) +
  scale_y_continuous(expand = expansion(c(0, 0.1)))+
  # facet_wrap(~experiment) +
  labs(title = "dropout",
       x = "",
       fill = "Compound",
       y = "Concentration [µM]")

dropout_MBOA_met
dropout_MBOA_met_export <- dropout_MBOA_met +
  scale_x_discrete(labels= rev(c("NBC", "LMB2", "LBA21", "LMG1", "LMN1", "LMX9", "LMX9231", "LST17", "metSC")))

# ggsave(plot = dropout_MBOA_met_export, filename = "Fig_2_dropout_MBOA_met.svg", path = "Output", width = 8.8, height = 9, dpi = 300, scale = 1, units = "cm")
```

```{r}
sessionInfo()
```


